---
title: "Opioid treatment checkins"
author: "Andrew Ba Tran"
date: "October 4, 2016"
output: html_document
---

Questions to ask:

- Which towns had the most admissions per capita?
- Urban versus rural versus mixed admissions?
Is there a correlation between admissions and overdose deaths? (By year)

Where do people tend to die (urban, rural, mixed)

distance to clinics, pharmacies, hospitals

correlation between deaths and admissions to these distances?



```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ctnamecleaner)
library(ggplot2)
library(extrafont)
library(scales)
library(rgdal)
require(maptools)
library(stringr)
library(knitr)
library(lubridate)
library("censusapi")
source("keys.R")

treat <- read.csv("https://data.ct.gov/api/views/4pv7-jhxb/rows.csv?accessType=DOWNLOAD", stringsAsFactors=F)

```

```{r total_town_year, message=F, warning=F, fig.width=8, fig.height=35}

treat$Admissions[is.na(treat$Admissions)] <- 0

gg <- ggplot(treat, aes(y=Admissions, x=FiscalYear))
gg <- gg + geom_bar(stat="identity")
gg <- gg + facet_wrap(~Town, ncol=4, scale="free_x")
#gg <- gg + facet_wrap(~state, scale="free", ncol=5)
#gg <- gg + scale_x_continuous(limits=c(2012,2016), breaks=c(2012,2016),
#        labels=c("2012", "2016"))
gg <- gg + labs(x=NULL, y=NULL, 
                title="Opioid Related Treatment Admissions by Town",
                subtitle="",
                caption="Department of Mental Health and Addiction Services")
gg <- gg + theme_minimal(base_family="Lato Regular")
gg <- gg + theme(panel.grid.major.y=element_blank())
gg <- gg + theme(panel.grid.minor=element_blank())
gg <- gg + theme(plot.title=element_text(family="Lato Black"))
gg


# gg <- ggplot(treat, aes(y=Unduplicated.Clients, x=FiscalYear))
# gg <- gg + geom_bar(stat="identity")
# gg <- gg + facet_wrap(~Town, ncol=4, scale="free_x")
# #gg <- gg + facet_wrap(~state, scale="free", ncol=5)
# #gg <- gg + scale_x_continuous(limits=c(2012,2016), breaks=c(2012,2016),
# #        labels=c("2012", "2016"))
# gg <- gg + labs(x=NULL, y=NULL, 
#                 title="Opioid Related Treatment Admissions (unduplicated clients) by Town",
#                 subtitle="",
#                 caption="Department of Mental Health and Addiction Services")
# gg <- gg + theme_minimal(base_family="Lato Regular")
# gg <- gg + theme(panel.grid.major.y=element_blank())
# gg <- gg + theme(panel.grid.minor=element_blank())
# gg <- gg + theme(plot.title=element_text(family="Lato Black"))
# gg
```

```{r total_town_year_per_capita, message=F, warning=F, fig.width=8, fig.height=35}

treat$Admissions[is.na(treat$Admissions)] <- 0
treat <- ctpopulator(Town, treat)
treat$per_capita_a <- round(treat$Admissions/treat$pop2013*1000, 2)

gg <- ggplot(treat, aes(y=per_capita_a, x=FiscalYear))
gg <- gg + geom_bar(stat="identity")
gg <- gg + facet_wrap(~Town, ncol=4, scale="free_x")
#gg <- gg + facet_wrap(~state, scale="free", ncol=5)
#gg <- gg + scale_x_continuous(limits=c(2012,2016), breaks=c(2012,2016),
#        labels=c("2012", "2016"))
gg <- gg + labs(x=NULL, y=NULL, 
                title="Opioid Related Treatment Admissions per 1,000 residents by Town",
                subtitle="",
                caption="Department of Mental Health and Addiction Services")
gg <- gg + theme_minimal(base_family="Lato Regular")
gg <- gg + theme(panel.grid.major.y=element_blank())
gg <- gg + theme(panel.grid.minor=element_blank())
gg <- gg + theme(plot.title=element_text(family="Lato Black"))
gg
```

# Mapped

```{r map1, message=F, warning=F, fig.width=9, fig.height=9}

town_shape <- readOGR(dsn="maps", layer="ctgeo")
town_shape_df <- fortify(town_shape, region="NAME10")

names(treat)[names(treat) == 'Town'] <- 'id'
treat$id <- str_to_title(treat$id)
town_map <- left_join(town_shape_df, treat)

gg <- ggplot(town_map, aes(long,lat, group=group, fill=per_capita_a)) 
gg <- gg +  geom_polygon()
gg <- gg +  geom_path(color = "grey73")
gg <- gg +  coord_equal()
gg <- gg + facet_wrap(~FiscalYear, ncol=2)
gg <- gg + scale_fill_gradient(low="grey73", high="blue") 
gg <- gg + labs(x=NULL, y=NULL, title="Opioid Related Treatment Admissions by town",
                subtitle="Per 1,000 residents",
                caption="SOURCE: Department of Mental Health and Addiction Services\nAndrew Ba Tran/TrendCT.org")
gg <- gg + theme_bw(base_family="Lato Regular")
gg <- gg + theme(text = element_text(size=16))
gg <- gg + theme(panel.grid.major=element_blank())
gg <- gg + theme(panel.grid.minor=element_blank())
gg <- gg + theme(panel.border=element_blank())
gg <- gg + theme(axis.ticks=element_blank())
gg <- gg + theme(axis.text.x=element_blank())
gg <- gg + theme(axis.text.y=element_blank())
gg <- gg + theme(plot.title=element_text(face="bold", family="Lato Black", size=22))
gg <- gg + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(b=12)))
gg <- gg + theme(plot.caption=element_text(size=12, margin=margin(t=10, r=80), color="#7a7d7e"))
gg <- gg + theme(plot.margin = unit(c(1, 1, 1, 1), "lines"))

gg

```

# Towns with most check-ins per capita

```{r table1}
tab1 <- treat %>%
  select(id, FiscalYear, per_capita_a) %>%
  spread(FiscalYear, per_capita_a) %>%
  arrange(-`2016`)
  kable(head(tab1, 10))
```

# Check-ins by town type

```{r town_type1, fig.width=9, fig.height=5, warning=F, message=F} 
source("urban_rural_mixed.R")

town_count <- town_count[c("NAME10", "Type", "perc_urban")]
colnames(town_count) <- c("id", "town_type", "perc_urban")

tab2 <- treat %>%
  select(id, FiscalYear, Admissions, pop2013, per_capita_a)

tab2_joined <- left_join(tab2, town_count)

gg <- ggplot(tab2_joined, aes(x=town_type, y=Admissions,fill=town_type))
gg <- gg +  geom_boxplot()
gg <- gg +  facet_wrap(~FiscalYear, ncol=5)
gg <- gg + labs(x=NULL, y="Admissions", title="Overall opioid-related checkins in Connecticut",
                subtitle="Per 1,000 residents",
                caption="SOURCE: Department of Mental Health and Addiction Services\nAndrew Ba Tran/TrendCT.org")
gg <- gg + theme_bw(base_family="Lato Regular")
gg <- gg + theme(text = element_text(size=16))
gg <- gg + theme(panel.grid.major=element_blank())
gg <- gg + theme(panel.grid.minor=element_blank())
gg <- gg + theme(panel.border=element_blank())
gg <- gg +  theme(axis.text.x=element_text(angle=90, vjust=0.4,hjust=1))
#gg <- gg + theme(axis.ticks=element_blank())
#gg <- gg + theme(axis.text.y=element_blank())
gg <- gg + theme(plot.title=element_text(face="bold", family="Lato Black", size=22))
gg <- gg + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(b=12)))
gg <- gg + theme(plot.caption=element_text(size=12, margin=margin(t=10, r=80), color="#7a7d7e"))
gg <- gg + theme(plot.margin = unit(c(1, 1, 1, 1), "lines"))

gg

p <- ggplot(tab2_joined, aes(x=town_type, y=per_capita_a,fill=town_type))
gg <- gg +  geom_boxplot()
gg <- gg +  facet_grid(.~FiscalYear)
gg <- gg + labs(x=NULL, y="Admissions per capita", title="Overall opioid-related checkins in Connecticut",
                subtitle="Per 1,000 residents",
                caption="SOURCE: Department of Mental Health and Addiction Services\nAndrew Ba Tran/TrendCT.org")
gg <- gg + theme_bw(base_family="Lato Regular")
gg <- gg + theme(text = element_text(size=16))
gg <- gg + theme(panel.grid.major=element_blank())
gg <- gg + theme(panel.grid.minor=element_blank())
gg <- gg + theme(panel.border=element_blank())
gg <- gg + theme(axis.text.x=element_text(angle=90, vjust=0.4,hjust=1))
#gg <- gg + theme(axis.ticks=element_blank())
#gg <- gg + theme(axis.text.y=element_blank())
gg <- gg + theme(plot.title=element_text(face="bold", family="Lato Black", size=22))
gg <- gg + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(b=12)))
gg <- gg + theme(plot.caption=element_text(size=12, margin=margin(t=10, r=80), color="#7a7d7e"))
gg <- gg + theme(plot.margin = unit(c(1, 1, 1, 1), "lines"))

gg
```

## Looking at deaths

Prepping the data
```{r deaths, message=F, warning=F, fig.width=8, fig.height=30}
drug_deaths <- read.csv("https://data.ct.gov/api/views/rybz-nyjw/rows.csv?accessType=DOWNLOAD", stringsAsFactors = F)

# Cleaning up race
drug_deaths$R <- "Unknown"
drug_deaths$R <- ifelse(drug_deaths$Race=="Asian Indian", "Asian", drug_deaths$R)
drug_deaths$R <- ifelse(drug_deaths$Race=="Asian, Other", "Asian", drug_deaths$R)
drug_deaths$R <- ifelse(drug_deaths$Race=="Chinese", "Asian", drug_deaths$R)
drug_deaths$R <- ifelse(drug_deaths$Race=="Hispanic, Black", "Hispanic", drug_deaths$R)
drug_deaths$R <- ifelse(drug_deaths$Race=="Hispanic, White", "Hispanic", drug_deaths$R)
drug_deaths$R <- ifelse(drug_deaths$Race=="Hispanic, Black", "Hispanic", drug_deaths$R)
drug_deaths$R <- ifelse(drug_deaths$Race=="Native American, Other", "Other", drug_deaths$R)
drug_deaths$R <- ifelse(drug_deaths$Race=="", "Unknown", drug_deaths$R)
drug_deaths$R <- ifelse(drug_deaths$Race=="Asian", "Asian", drug_deaths$R)
drug_deaths$R <- ifelse(drug_deaths$Race=="Black", "Black", drug_deaths$R)
drug_deaths$R <- ifelse(drug_deaths$Race=="Hispanic", "Hispanic", drug_deaths$R)
drug_deaths$R <- ifelse(drug_deaths$Race=="White", "White", drug_deaths$R)

# Cleaning up death and residence city locations

drug_deaths <- ctnamecleaner(Death.City, drug_deaths)
names(drug_deaths)[names(drug_deaths) == 'real.town.name'] <- 'town.of.death'
drug_deaths$Residence.City <- gsub("  ", " ", drug_deaths$Residence.City)
drug_deaths <- ctnamecleaner(Residence.City, drug_deaths)
names(drug_deaths)[names(drug_deaths) == 'real.town.name'] <- 'town.of.residence'
drug_deaths$residence.and.death <- ""
drug_deaths$residence.and.death <- ifelse(drug_deaths$town.of.death == drug_deaths$town.of.residence, "Same town of death and residence", "Different town of death than residence")
drug_deaths$residence.and.death <- ifelse(is.na(drug_deaths$town.of.death), "Town of death unknown", drug_deaths$residence.and.death)
drug_deaths$residence.and.death <- ifelse(is.na(drug_deaths$town.of.residence), "Residence unknown", drug_deaths$residence.and.death)
drug_deaths$residence.and.death <- ifelse((is.na(drug_deaths$town.of.residence) & (drug_deaths$Residence.State!="CT")), "Out of town resident, local death", drug_deaths$residence.and.death)
drug_deaths$residence.and.death <- ifelse((is.na(drug_deaths$town.of.residence) & (drug_deaths$Residence.City=="")), "Unknown residence", drug_deaths$residence.and.death)

drug_deaths$date.fixed <- mdy(drug_deaths$Date)
drug_deaths$Year <- year(drug_deaths$date.fixed)
drug_deaths$row <- row.names(drug_deaths)

# fentanyl only
new_fentanyl <- drug_deaths %>%
  mutate(cause=str_to_lower(drug_deaths$ImmediateCauseA)) %>%
  filter(grepl("fent", cause) | Fentanyl=="Y" )

new_fent <-  select(new_fentanyl, row)
new_fent$update_fent <- "Yes"
new_fent <- as.data.frame(new_fent)
drug_deaths <- left_join(drug_deaths, new_fent)
  
# heroin only

new_heroin <- drug_deaths %>%
  mutate(cause=str_to_lower(drug_deaths$ImmediateCauseA)) %>%
  filter(grepl("heroin", cause) | Heroin=="Y" )

new_her <- select(new_heroin, row)
new_her$update_heroin <- "Yes"
new_her <- as.data.frame(new_her)
drug_deaths <- left_join(drug_deaths, new_her)

# prescription opioids

new_presc <- drug_deaths %>%
  mutate(cause=str_to_lower(drug_deaths$ImmediateCauseA)) %>%
  filter(grepl("oxy", cause) | Oxycodone=="Y" | Oxymorphone=="Y" |
           grepl("hydro", cause) | Hydrocodone=="Y" |
           grepl("tramad", cause) | Tramad=="Y" |
           grepl("morphine", cause) | Morphine..not.heroin.=="Y" )


new_pre <- select(new_presc, row)
new_pre$update_presc <- "Yes"
new_pre <- as.data.frame(new_pre)
drug_deaths <- left_join(drug_deaths, new_pre)

#leftover <- subset(drug_deaths, is.na(update_fent) & is.na(update_heroin) & is.na(update_presc))
#ggplot(leftover, aes(Year)) + geom_bar()

# There were about 461 cases that did not involve overdoses of prescription opioids, heroin, or fentanyl

# OK, checkins and deaths by year and town


# Prepping town and year dataframe

towns_only <- data.frame(town_count[,1])
colnames(towns_only) <- "town.of.residence"
towns_only$Year <- 0

for (i in 2012:2016) {
  towns_only$Year <- i
  if (i == 2012) {
    mega_towns <- towns_only
  } else {
    mega_towns <- rbind(mega_towns, towns_only)
  }
}

# First, deaths by town (just opioids)

opioid_deaths <- drug_deaths %>%
  filter(update_presc=="Yes") %>%
  group_by(town.of.residence, Year) %>%
  summarise(overdose.deaths=n())
# 
checkins <- tab2
colnames(checkins) <- c("town.of.residence", "Year", "admissions", "pop", "per_capita_a")

test <- left_join(mega_towns, checkins)
test <- left_join(test, opioid_deaths)

cor(test$admissions, test$overdose.deaths, use="complete.obs")
cor(test$per_capita_a, test$overdose.deaths, use="complete.obs")
test2 <- test
test2$per_capita_o <- round(test2$overdose.deaths/test2$pop*10000,2)
test2$overdose.deaths[is.na(test2$overdose.deaths)] <- 0
cor(test2$admissions, test2$overdose.deaths, use="complete.obs")
cor(test2$per_capita_a, test2$overdose.deaths, use="complete.obs")
cor(test2$per_capita_a, test2$per_capita_o, use="complete.obs")

test3 <- test2 %>%
  select(town.of.residence, Year, per_capita_a, per_capita_o) %>%
  gather("Type", "per.capita", 3:4)

gg <- ggplot(test3, aes(x=Year, y=per.capita, group=Type, color=Type))
gg <- gg + geom_line()
gg <- gg + facet_wrap(~town.of.residence, scales="free", ncol=5)
gg

test4 <- test2 %>%
  select(town.of.residence, Year, admissions, overdose.deaths) %>%
  gather("Type", "Count", 3:4)

gg <- ggplot(test4, aes(x=Year, y=Count, group=Type, color=Type))
gg <- gg + geom_line()
gg <- gg + facet_wrap(~town.of.residence, scales="free", ncol=5)
gg
```


```{r town_totals}

opioid_deaths2 <- drug_deaths %>%
  filter(update_presc=="Yes") %>%
  group_by(town.of.residence) %>%
  summarise(overdose.deaths=n())
# 
checkins <- tab2
colnames(checkins) <- c("town.of.residence", "Year", "admissions", "pop", "per_capita_a")

checkins2 <- checkins %>%
  group_by(town.of.residence) %>%
  summarise(admissions=sum(admissions))

town_totals <- left_join(towns_only, opioid_deaths2)
town_totals <- left_join(town_totals, checkins2)


cor(town_totals$admissions, town_totals$overdose.deaths, use="complete.obs")
town_totals <- ctpopulator(town.of.residence, town_totals)

town_totals$per_cap_o <- round(town_totals$overdose.deaths/town_totals$pop2013*1000,2)
town_totals$per_cap_a <- round(town_totals$admissions/town_totals$pop2013*1000,2)
cor(town_totals$per_cap_a, town_totals$per_cap_o, use="complete.obs")

gg <- ggplot(town_totals, aes(x=overdose.deaths, y=admissions))
gg <- gg + geom_point()
gg <- gg + labs(x="Overdoses", y="Admissions", 
                title="Overdose deaths compared to clinical admissions by town",
                subtitle="",
                caption="Department of Mental Health and Addiction Services")
gg

gg <- ggplot(town_totals, aes(x=per_cap_o, y=per_cap_a))
gg <- gg + geom_point()
gg <- gg + labs(x="Overdoses per capita", y="Admissions per capita", 
                title="Overdose deaths compared to clinical admissions by town (per 1,000 residents)",
                subtitle="",
                caption="Department of Mental Health and Addiction Services")
gg

```


# Income


```{r towns_income_admissions}
data2014 <- getCensus(name="acs5", 
                      vintage=2014,
                      key=census_key, 
                      vars=c("NAME", "B19013_001E"), 
                      region="county subdivision:*",  regionin="state:09")
data2014$town.of.residence <- gsub(" town.*", "", data2014$NAME)
data2014 <- subset(data2014, !is.na(B19013_001E))
data2014 <- data2014[,5:6]

colnames(data2014) <- c("median.income", "town.of.residence")

town_totals2 <- town_totals 
town_totals2$town.of.residence <- str_to_title(town_totals2$town.of.residence)
town_totals2 <- left_join(town_totals2, data2014)

cor(town_totals2$per_cap_a, town_totals2$median.income, use="complete.obs")

gg <- ggplot(town_totals2, aes(x=median.income, y=per_cap_a))
gg <- gg + geom_point()
gg <- gg + labs(x="Median income", y="Admissions per capita", 
                title="Clinical admissions by town (per 1,000 residents) compared to median income",
                subtitle="",
                caption="Department of Mental Health and Addiction Services")
gg

cor(town_totals2$per_cap_o, town_totals2$median.income, use="complete.obs")

gg <- ggplot(town_totals2, aes(x=median.income, y=per_cap_o))
gg <- gg + geom_point()
gg <- gg + labs(x="Median income", y="Overdoses per capita", 
                title="Opioid overdoses by town (per 1,000 residents) compared to median income",
                subtitle="",
                caption="Department of Mental Health and Addiction Services")
gg
```


```{r box3, fig.width=9, fig.height=5, warning=F, message=F}
tab3_joined <- town_count
colnames(tab3_joined) <- c("town.of.residence", "town_type", "perc_urban")
tab3_joined <- left_join(test2, tab3_joined)


gg <- ggplot(tab3_joined, aes(x=town_type, y=per_capita_o, group=town_type, fill=town_type))
gg <- gg +  geom_boxplot()
gg <- gg +  facet_grid(.~Year)
gg <- gg + labs(x=NULL, y="Overdoses per capita", title="Adjusted opioid-related overdoses in Connecticut",
                subtitle="Per 1,000 residents",
                caption="SOURCE: Department of Mental Health and Addiction Services\nAndrew Ba Tran/TrendCT.org")
gg <- gg + theme_bw(base_family="Lato Regular")
gg <- gg + theme(text = element_text(size=16))
gg <- gg + theme(panel.grid.major=element_blank())
gg <- gg + theme(panel.grid.minor=element_blank())
gg <- gg + theme(panel.border=element_blank())
gg <- gg + theme(axis.text.x=element_text(angle=90, vjust=0.4,hjust=1))
#gg <- gg + theme(axis.ticks=element_blank())
#gg <- gg + theme(axis.text.y=element_blank())
gg <- gg + theme(plot.title=element_text(face="bold", family="Lato Black", size=22))
gg <- gg + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(b=12)))
gg <- gg + theme(plot.caption=element_text(size=12, margin=margin(t=10, r=80), color="#7a7d7e"))
gg <- gg + theme(plot.margin = unit(c(1, 1, 1, 1), "lines"))
gg
```

```{r scatter2, fig.width=9, fig.height=12}
gg <- ggplot(tab3_joined, aes(x=per_capita_a, y=per_capita_o, group=town_type, color=town_type))
gg <- gg +  geom_point()
gg <- gg +  facet_wrap(~Year, ncol=2)
gg <- gg + labs(x="Admissions per capita", y="Overdoses per capita", title="Opioid-related overdoses compared to treatment admissions",
                subtitle="Per 1,000 residents",
                caption="SOURCE: Department of Mental Health and Addiction Services\nAndrew Ba Tran/TrendCT.org")
gg <- gg + theme_bw(base_family="Lato Regular")
gg <- gg + theme(text = element_text(size=16))
gg <- gg + theme(panel.grid.major=element_blank())
gg <- gg + theme(panel.grid.minor=element_blank())
gg <- gg + theme(panel.border=element_blank())
gg <- gg + theme(axis.text.x=element_text(angle=90, vjust=0.4,hjust=1))
#gg <- gg + theme(axis.ticks=element_blank())
#gg <- gg + theme(axis.text.y=element_blank())
gg <- gg + theme(plot.title=element_text(face="bold", family="Lato Black", size=22))
gg <- gg + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(b=12)))
gg <- gg + theme(plot.caption=element_text(size=12, margin=margin(t=10, r=80), color="#7a7d7e"))
gg <- gg + theme(plot.margin = unit(c(1, 1, 1, 1), "lines"))
gg

```
